---
title: "Week 4 peer-reviewed Prediction Assignment Writeup"
author: "Tingfen Yan"
date: "9/2/2017"
output: html_document
---
##Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

##Data

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

##load library
```{r message=FALSE,warning=FALSE}
library(caret)
library(randomForest)
library(gbm)
```
##load and preprocess data
```{r echo=TRUE}
set.seed(2017-09-02)
##loading the dataset
trainUrl = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training = read.csv(url(trainUrl), na.strings = c("NA", "#DIV/0!", ""))
testing = read.csv(url(testUrl), na.strings = c("NA", "#DIV/0!", ""))
##remove the columns are not related to predition variables required
training1 = training[,-c(1:7)]
testing1 = testing[, -c(1:7)]
##Remove those variables with NA value for training and testing set
training1=training1[, colSums(is.na(training1))==0]
testing1=testing1[, colSums(is.na(testing1))==0]

dim(training1); dim(testing1)
```

```{r}
##the predictor variables showed in both training and testing dataset
int =intersect(names(training1), names(testing1))
length(int)
```
Except variable classe in training and problem_id in testing, the rest 52 predition variables are common for both dataset.

```{r}
training1$classe <- as.factor(training1$classe)
##partition training1 dataset into training and probing two set by outcome classe
inTrain=createDataPartition(training1$classe, p=0.6, list=FALSE)
training2=training1[inTrain,]
probing=training1[-inTrain,]
dim(training2); dim(probing)
```
###Prediction model1 random forest
```{r}
modFit=train(classe ~ ., method="rf", data=training2)
pred_probing=predict(modFit, probing)
##accuracy from random forest
confusionMatrix(pred_probing, probing[, "classe"])$overall[1]
```

###Prediction model2 boosting machine model
```{r}
modgbm=train(classe ~ ., method="gbm", data=training2,verbose=FALSE)
predgm=predict(modgbm, probing)
##accuracy from boosting
confusionMatrix(predgm,probing$classe)$overall[1]
```
Both Random forest and boosting are good. Random forest is more accurate than boosting. I will use random for further analysis
```{r}
modFit$finalModel
```
The error rate is 0.79% less than 1%. 
## apply the Random forest model to testing set
```{r}
pred_testing=predict(modFit,testing1[,-ncol(testing1)])
pred_testing
```






